---
title: "Sumif / Sumifs"
author: "Michael Gordon"
date: "9 February 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How to run sumif/sumifs functions in R

As with most data manipulation we will be using the dplyr package, which is part of the [tidyverse](https://blog.rstudio.org/2016/09/15/tidyverse-1-0-0/). The key function for us is summarise(). The summarise() function will create summary statistics for a given column in the data frame such as finding the mean or summing. 
 
I will use the package dplyr and will create a data set below.
```{r cars}
# Load Package
library(dplyr)
# Create Dataframe
Area <- c("Auckland","Auckland","Auckland","Waikato","Waikato","Waikato","Wellington","Wellington","Wellington","Rest of New Zealand","Rest of New Zealand","Rest of New Zealand")
Industry <- c("Primary","Secondary","Tertiary","Primary","Secondary","Tertiary","Primary","Secondary","Tertiary","Primary","Secondary","Tertiary")
GDP <- c(56,200,354,150,75,79,45,67,197,300,420,77)
GEOs <- c(1001,2678,5080,506,768,923,1076,111,234,678,299,651)
MECs <- c(3000,7689,15903,1589,2134,2765,3200,345,700,1901,1100,1789)
data <- data.frame(Area,Industry,GDP,GEOs,MECs)

# View Data Frame
head(data,12)
```
 
Lets try summarise a column - using sum() inside summarise

```{r}
data %>% summarise(sum(GDP))
```
 
A major strength of dplyr is the ability to group the data by a variable or variables and then operate on the data "by group". Here we shall sum by MECs by Industry. This is the equivalent of Sumifs.

```{r}
data %>% group_by(Industry) %>%  summarise(Sum_of_MECs = sum(MECs))
```
 
To sumif by a single condition we use filter first. Say we are only interested in MECs in primary industry
```{r}
data %>% filter(Industry == "Primary") %>% group_by(Industry) %>%  summarise(Sum_of_MECs = sum(MECs))
```
  
  
To run a sumifs we simply group_by multiple variables. 
```{r}
data %>% group_by(Industry,Area) %>%  summarise(Sum_of_MECs = sum(MECs))
```


